Supplementary Material

### 2024-04-19 products/manuscript/supplement/Supplementary.qmd
Project Part 5
Completed part 5 of the project (might review and do a few edits later)

---
title: "Supplementary Material"
subtitle: "Project: Part 3"
author: Erick E. Mollinedo
date: '`r format(Sys.Date(), "%B %d, %Y")`'
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
---

```{r include=FALSE}
library(tidyverse)
library(tidymodels)
library(here)
library(readr)
```

@@

@fig-conc shows the summary of concentrations for each chemical species categorized by study arm. It is observed a similar pattern observed from the categorization by type of fuel, where Mg, Mn, Ca, K, BC, Ti, Si and S have statistically significant differences in concentrations (shown as ****).

![Chemical species concentrations by study arm](../../../results/figures/boxplot-arm.jpg){#fig-conc}

@@

![Carbon Monoxide concentrations by type of fuel](../../../results/figures/co.jpg){#fig-co}

![PM2.5 concentrations by type of fuel](../../../results/figures/pm25.jpg){#fig-pm}

The following tables show the results when conducting the generalized linear models using a single or multiple variables for the chemical species that can be interpreted by sources of exposure.

```{r echo=FALSE}
#Open the hapin_merged df

hapin_merged <- read_csv(here("data", "processed-data", "hapin-merged.csv"))

# Create a generalized linear regression model specification using gamma distribution (This will be used for all the models)
model_spec <- linear_reg(mode = "regression") %>%
  set_engine("glm", family = Gamma(link = "log"))

###########To predict the BC variable##############
# Create the recipe for preprocessing
recipeBC_red <- recipe(BC ~ kerosene + stove_other + fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowBC_red <- workflow() %>%
  add_recipe(recipeBC_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitBC_red <- workflowBC_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsBC_red <- fitBC_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsBC_red)

###########To predict the Ca variable##############
# Create the recipe for preprocessing
recipeCa_red <- recipe(Ca ~ fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowCa_red <- workflow() %>%
  add_recipe(recipeCa_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitCa_red <- workflowCa_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsCa_red <- fitCa_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsCa_red)

###########To predict the K variable##############
# Create the recipe for preprocessing
recipeK_red <- recipe(K ~ smoke + stove_other + fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowK_red <- workflow() %>%
  add_recipe(recipeK_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitK_red <- workflowK_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsK_red <- fitK_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsK_red)

###########To predict the Mg variable##############
# Create the recipe for preprocessing
recipeMg_red <- recipe(Mg ~ kerosene + fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowMg_red <- workflow() %>%
  add_recipe(recipeMg_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitMg_red <- workflowMg_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsMg_red <- fitMg_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsMg_red)

###########To predict the Mn variable##############
# Create the recipe for preprocessing
recipeMn_red <- recipe(Mn ~ fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowMn_red <- workflow() %>%
  add_recipe(recipeMn_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitMn_red <- workflowMn_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsMn_red <- fitMn_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsMn_red)

###########To predict the S variable##############
# Create the recipe for preprocessing
recipeS_red <- recipe(S ~ fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowS_red <- workflow() %>%
  add_recipe(recipeS_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitS_red <- workflowS_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsS_red <- fitS_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsS_red)

###########To predict the Fe variable##############
# Create the recipe for preprocessing
recipeFe_red <- recipe(Fe ~ stove, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowFe_red <- workflow() %>%
  add_recipe(recipeFe_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitFe_red <- workflowFe_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsFe_red <- fitFe_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsFe_red)

###########To predict the Si variable##############
# Create the recipe for preprocessing
recipeSi_red <- recipe(Si ~ fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowSi_red <- workflow() %>%
  add_recipe(recipeSi_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitSi_red <- workflowSi_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsSi_red <- fitSi_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsSi_red)

###########To predict the Ti variable##############
# Create the recipe for preprocessing
recipeTi_red <- recipe(Ti ~ fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowTi_red <- workflow() %>%
  add_recipe(recipeTi_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitTi_red <- workflowTi_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsTi_red <- fitTi_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsTi_red)
```

>>>>

---
title: "Supplementary Material"
author: Erick E. Mollinedo
date: '`r format(Sys.Date(), "%B %d, %Y")`'
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
---

```{r include=FALSE}
library(tidyverse)
library(tidymodels)
library(here)
library(readr)
library(gt)
```

@@

@fig-conc shows the summary of concentrations for each chemical species categorized by study arm. It is observed a similar pattern observed from the categorization by type of fuel, where Mg, Mn, Ca, K, BC, Ti, Si and S have statistically significant differences in concentrations (shown as \*\*\*\*).

![Chemical species concentrations by study arm](../../../results/figures/boxplot-arm.png){#fig-conc}

@@

![Carbon Monoxide concentrations by type of fuel](../../../results/figures/co.png){#fig-co}

![PM2.5 concentrations by type of fuel](../../../results/figures/pm25.png){#fig-pm}

The following tables show the model summaries for the final models selected. In this case the linear regression model for the PM~2.5~ mass concentrations and the gamma GLM model for the BC mass concentrations.

```{r echo=FALSE}
#Load the `hapin_merged` dataframe
hapin_merged <- read_rds(here("data", "processed-data", "hapin-final.rds"))

# Split the data into training and testing sets
rngseed = 1234
set.seed(rngseed) # for reproducibility
data_split <- initial_split(hapin_merged, prop = 0.8)
train_data <- training(data_split)
test_data <- testing(data_split)

#Create formulas
pm25formula <- pm25 ~ Mg + Al + Si + S + K + Ca + Ti + Mn + Fe + Zn + BC
bcformula <- BC ~ smoke + coil + trash + kerosene + smoky + crop + stove_other + fueltype

#Set model specifications
lin_mod <- linear_reg() %>% set_engine("lm") #Linear
glm_mod <- linear_reg(mode = "regression") %>% #GLM gamma
  set_engine("glm", family = Gamma(link = "log"))

## ----Final model PM2.5: Linear regression---- ##
linpm_wflow <- workflow() %>% #Workflow
	add_model(lin_mod) %>% 
	add_formula(pm25formula)

linpm_fit <- linpm_wflow %>% fit(data = train_data) #Fit the model

## ----Final model BC: GLM Gamma---- ##
glmbc_wflow <- workflow() %>% #Workflow
  add_model(glm_mod) %>% 
  add_formula(bcformula)

glmbc_fit <- glmbc_wflow %>% fit(data = train_data) #Fit the model

# Extract and print the results of the models
results_pmlin <- linpm_fit %>% #PM2.5 Linear
  extract_fit_parsnip() %>%
  tidy()

results_bcglm <- glmbc_fit %>% #BC GLM
  extract_fit_parsnip() %>%
  tidy()

#Print the results
print(results_pmlin)
print(results_bcglm)
```

The following figures illustrate the observed vs predicted values plots for both PM~2.5~ and BC models, and the residual plots.

![Observed vs Predicted values of the PM2.5 models](../../../results/figures/obs-pred_pm25.png)

![PM2.5 Models residuals](../../../results/figures/residuals-pm25.png)

![Observed vs Predicted values of the BC models](../../../results/figures/obs-pred_bc.png)

![BC models residuals](../../../results/figures/residuals-bc.png)
